---
title: "Audit everything"
subtitle: "Registry-wide replication health check"
author: "replicateEverything"
date: last-modified
params:
  patience: 20
  install_deps: false
  dois: !expr NULL
execute:
  echo: false
  warning: false
  message: false
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    theme: flatly
---

```{r setup}
#| include: false
knitr::opts_chunk$set(comment = NA)

if (!requireNamespace("replicateEverything", quietly = TRUE)) {
  stop(
    "Install replicateEverything before rendering this report, e.g.\n",
    "  remotes::install_github('replicate-anything/replicateEverything')",
    call. = FALSE
  )
}

library(replicateEverything)

patience <- if (is.null(params$patience)) 20 else params$patience
install_deps <- isTRUE(params$install_deps)
dois <- params$dois

audit <- audit_everything(
  patience = patience,
  dois = dois,
  install_deps = install_deps,
  verbose = TRUE
)

results <- audit$results
summary <- audit$summary
```

This report runs [audit_everything()] across the replication registry. Each table and figure is replicated in every listed engine (R and Stata where available). A run is halted after **`r patience`** seconds per object; failures are recorded and the audit continues.

**Started:** `r format(audit$started_at, "%Y-%m-%d %H:%M:%S %Z")`  
**Finished:** `r format(audit$finished_at, "%Y-%m-%d %H:%M:%S %Z")`

## Summary

| Metric | Count |
|:--|--:|
| Studies audited | `r summary$studies` |
| Replication runs | `r summary$runs` |
| Successful | `r summary$success` |
| Failed | `r summary$failed` |
| Timed out | `r summary$timed_out` |

```{r summary-bar}
#| echo: false
if (summary$runs > 0) {
  pct <- round(100 * summary$success / summary$runs, 1)
  cat(sprintf("**Pass rate:** %s%%\n\n", pct))
}
```

## Results by study

```{r by-study}
#| results: asis
if (nrow(results) == 0) {
  cat("_No replication runs were recorded._\n")
} else {
  studies <- unique(results$title)
  for (study in studies) {
    cat("\n\n### ", study, "\n\n", sep = "")
    sub <- results[results$title == study, , drop = FALSE]
    sub$status <- ifelse(
      sub$success,
      "OK",
      ifelse(sub$timed_out, "Timed out", "Failed")
    )
    sub$seconds <- ifelse(is.na(sub$seconds), NA, round(sub$seconds, 2))
    show <- sub[, c(
      "object_label", "object", "type", "engine",
      "status", "seconds", "error_snippet"
    )]
    names(show) <- c(
      "Object", "ID", "Type", "Engine",
      "Status", "Seconds", "Error"
    )
    print(knitr::kable(show, format = "html", na = ""))
  }
}
```

## Failed runs (concise)

```{r failures}
#| echo: false
fails <- results[!results$success, , drop = FALSE]
if (nrow(fails) == 0) {
  cat("_All recorded runs succeeded._\n")
} else {
  fails$seconds <- ifelse(is.na(fails$seconds), NA, round(fails$seconds, 2))
  fails$status <- ifelse(fails$timed_out, "Timed out", "Failed")
  show <- fails[, c(
    "title", "object_label", "object", "engine",
    "status", "seconds", "error_snippet"
  )]
  names(show) <- c(
    "Study", "Object", "ID", "Engine",
    "Status", "Seconds", "Error"
  )
  knitr::kable(show, format = "html", na = "")
}
```

## Re-run locally

```r
library(replicateEverything)
audit <- audit_everything(patience = 20)
print(audit)
```

Render this document from the package checkout:

```bash
quarto render audit_everything.qmd -P patience:30
```

Or a subset of studies:

```r
quarto::quarto_render(
  replicateEverything::audit_everything_qmd(),
  execute_params = list(
    patience = 20,
    dois = c("10.1177/00491241211036161")
  )
)
```
