---
title: "replicateEverything"
format: gfm
---

# replicateEverything

A platform for **reproducing academic research results programmatically and simply**.

The package allows users to reproduce figures and tables from academic papers using a standardized replication registry.

The goal is to make research **transparent, modular, and easily reproducible**.

---

## Overview

`replicateEverything` provides a simple interface to:

- search replication-ready papers
- retrieve replication repositories
- reproduce figures and tables
- run full paper replications

The system is built around a **registry architecture** where replication repositories are indexed and accessed dynamically.

---


## Installation
Install the development version from GitHub by using `remotes` and `devtools`.

```{r}
if (!requireNamespace("replicateEverything", quietly = TRUE)) {
  options(repos = c(CRAN = "https://cloud.r-project.org"))
  remotes::install_github("replicate-anything/replicateEverything")
}
```



## Load Package 
Load the package in your environment. 
```{r}
library(replicateEverything)
```

# How the Package Works
The package in this very early version allows users to search for papers in the repository, find the replication repository for a paper, talk look at the list of existing replications, reproduce a specific result (e.g. **Fig.1**), reproduce all outputs (i.e. **Fig.1**, **Fig.2**, **Table 1**, etc).  

## Search Papers 
To search papers, you should run the command below. You have to specify closely associated words with the paper. 
```{r}
# first option 
search_papers("minimum wage")

# second option 
search_papers("wage")

# third option
search_papers("employment")

```

You can use many **key words** to search from the **title** of a paper to see if it is in the directory. 

## Find Replication Repository 
The way the package is set up now, researchers will have to outline their Github repository is an aligned way to allows it to be merge into the [Registry](https://github.com/replicate-anything/registry). More details on this later. However, you can use the `find_repo` function and the paper's `DOI` to located and look up a paper's repository from the registry. 
```{r}
# find repo
find_repo("10.1257/aer.20221688")
```


## List Repository 
The `list_replications` function allows you to look up **structures** in an uploaded repository. It also allows you to see some important **$id* that are subsequently used in key replication functions in the package, for example `fig_1`.
```{r}
list_replications("10.1257/aer.20221688")
```

## Run Replication
The `run_replication` function allows you to run **single** replication for a specific table or figure. This must have already be specified in the root directory on Github. More details on this later when discussing directory. 
```{r}
run_replication(
  "10.1257/aer.20221688",
  "fig_1"
)
```

## Replicate Paper 
This is the **single most important function** in this `replicateEverything` package. It allows you to replicate an entire paper. With a single line of code, you can be able to generate all figures at one go. See how it works below:  
```{r}
replicate_paper("10.1257/aer.20221688")
```

# System Architecture
```{mermaid}

flowchart LR
  A[Registry] --> B(Replication Repositories)
  B --> C{replicateEverything}
```

## Registry 
The registry indexes all replication repositories. See an example entry:
```{}
#| include: false

    "doi": "10.1257/aer.20221688",
    "title": "Market Design and Moral Behavior",
    "authors": ["Bartling", "Fehr"],
    "year": 2024,
    "journal": "American Economic Review",
    "repo": "replicate-anything/aer-replications"

```

## Repositories Structures 
Every replication repository **MUST** follows a standardized structure. All data must be store in `processed` folder. All code must be stored in the `code` folder.
```{}
#| include: false

papers/
   DOI/
      replication.yml
      processed/
         fig_1.csv
         tab_1.csv
      code/
         fig_1.R
         tab_1.R
```

### Example 
See the example below for structure of a repo for a single figure. This should similarly work for a single table as well. The repo should have the data and code as seen below. 
```{}

papers/
   10.1257_aer.20221688/
      replication.yml
      processed/
         fig_1.csv
      code/
         fig_1.R

```


# Metadata Specification 
Each paper must include a metadata file. 
```{}
replication.yml
```

## Example
Ideally, make sure to include if a table has `dependencies`. 
```{YAML}
#| include: false

paper:
  title: Market Design and Moral Behavior
  authors:
    - Bartling, Björn
    - Fehr, Ernst
  year: 2024
  doi: 10.1257/aer.20221688
  journal: American Economic Review

replications:

  - id: fig_1
    type: figure
    description: Example figure
    data: processed/fig_1.csv
    code: code/fig_1.R

  - id: tab_1
    type: table
    description: Example table
    data: processed/tab_1.csv
    code: code/tab_1.R
    dependencies:
      - dplyr
      - gt

```


# Writing Replication Scripts 
Replications Scripts **must** define the `generate` function. For figure, use `generate_figure` and for table, use `generate_table`. 

## Figures
```{r}
#| include: false

generate_figure <- function(data){

  ggplot2::ggplot(data, ggplot2::aes(group,value)) +
    ggplot2::geom_col()

}

```


## Tables 
```{r}
#| include: false
generate_table <- function(data){

  dplyr::summarise(data, mean_value = mean(value))

}
```

This allows the the `package` to automatically: 
* download the datasets
* loads dependencies 
* sources the script 
* runs the function 

# Contributing to a Repository 
This is a very important aspect in our quest to create a reprocible research culture so please follow the instructions below keenly: 

1. Fork the replication repository

2. Add a new paper directory

3. Provide metadata and scripts

4. Submit a pull request


See example below: 
```{}
papers/10.xxxx_xxxx/
```

# Developer Workflow 
Start by cloning the repository: 

```{}
git clone https://github.com/replicate-anything/replicateEverything
```

Install the dependencies: 
```{r}
#| include: false
devtools::install()
```

Run package checks: 
```{r}
#| include: false
devtools::check()
```


# Project Status
This idea is conceived by [Macartan Humphreys](https://github.com/replicate-anything/registry), Director of the IPI Research Unit at [WZB](https://www.wzb.eu) in 2024. The project is under active development at the moment. All feedback are welcome. Feel free to [email me](mailto:vermon.washington@wzb.eu). 
