---
title: "Registry audit"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Registry audit}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
audit_rds <- system.file(
  "vignette-data", "audit_latest.rds",
  package = "replicateEverything"
)
if (!nzchar(audit_rds) || !file.exists(audit_rds)) {
  candidates <- c(
    "inst/vignette-data/audit_latest.rds",
    "../inst/vignette-data/audit_latest.rds"
  )
  found <- candidates[file.exists(candidates)]
  audit_rds <- if (length(found)) found[[1]] else ""
}
run_live <- identical(
  tolower(Sys.getenv("REPLICATE_AUDIT_LIVE", unset = "")),
  "true"
)
```

## What `audit_everything()` does

`audit_everything()` walks the [replication registry](https://github.com/replicate-anything/registry) and attempts **every table and figure** in **each listed engine** (R and Stata where both exist). It is meant as a health check: some failures are expected as studies, data, and dependencies change over time.

Key behaviour:

- **`patience`** (default `20` seconds; registry reports often use `60`) — each table or figure is halted after this audit cap; the audit **continues** with the next object. Timeout rows record `timeout_seconds` and an explicit “Timed out after N seconds (audit cap)” message.
- **Failures do not stop the run** — results are collected in a data frame.
- **Incomplete / unavailable steps are skipped** — yaml `incomplete: true` (including `requires_engine:` / `data_unavailable:` gaps) is **not** attempted. Rows are recorded with status **Skipped** and a reason, and are **not** counted as success or failure. Distinct from fail/timeout.
- **Report fields** — study, object id, engine, success, skipped flag, elapsed seconds, timeout_seconds (audit cap), timed-out flag, and a short error / skip-reason snippet.
- **Substantive checks** (default `substantive = TRUE`) — when a study defines `tests/substantive/<step_id>.R`, the audit compares replicated estimates to published benchmarks (see Fearon & Laitin `tab_1`). Failures appear as `[substantive]` in the printed summary.
- **Related studies** are compiled when the registry index is rebuilt ([refresh_registry()] / internal index build): upstream from `paper.related` / `paper.extends`, downstream by reversing those pointers. Use `summary(get_study(doi))` for a quick console view; the audit itself does not re-fetch related links per run.

```{r, eval = FALSE}
library(replicateEverything)

# Point at a local monorepo checkout (optional)
options(
  replicateEverything.registry_root = "/path/to/replicate_everything/registry",
  replicateEverything.study_folders_root = "/path/to/replicate_everything",
  replicateEverything.use_sibling_packages = TRUE
)

audit <- audit_everything(patience = 20)
print(audit)
```

## Incremental CSV job store

As of replicateEverything **0.7.40+**, registry audits no longer replace the whole portfolio snapshot on every call. When a registry root is set, each run **upserts** into a flat CSV and rebuilds derived files from the **full** store.

### Source of truth: `registry/audit_jobs.csv`

| File | Role |
|------|------|
| `audit_jobs.csv` | **Source of truth** — one row per job, keyed by **doi × object × engine** |
| `audit_summary.json` | Derived portfolio counts + `progress` (Shiny health bar) |
| `audit_latest.rds` | Derived full `audit_everything` object rebuilt from the CSV |

Upsert means: rows for jobs in this run are updated; every other study’s rows stay put. A one-DOI or one-collection audit therefore **does not wipe** the Shiny health bar or drop other studies from the derived summary.

### Maintainer verbs (0.7.41)

Four public maintainer verbs cover registration and audit:

| Verb | Role |
|------|------|
| [register_study()] | Validate + sync one study stub into the registry |
| [refresh_registry()] | **Light by default**: rebuild index + Shiny cache, seed CSV gaps, rebuild summary. Pass `audit = TRUE` or `audit = <dois>` for live runs |
| [audit_everything()] | Manual / heavy / Quarto live audit |
| [audit_report()] | **Read-only** portfolio health from CSV / summary (no writes) |

```{r, eval = FALSE}
# Light refresh (no live engines)
refresh_registry("registry")

# Light refresh + full live audit
refresh_registry("registry", audit = TRUE, patience = 20)

# Light refresh + subset live audit
refresh_registry("registry", audit = "10.1257/aer.91.5.1369")

# Read-only console / dataframe view
audit_report("registry")
```

### Full portfolio vs add / refresh a subset

```{r, eval = FALSE}
# Full registry (every index row)
audit <- audit_everything(patience = 20)

# Add or refresh only these DOIs / collections (upsert; others kept)
audit <- audit_everything(patience = 30, dois = "10.1257/aer.91.5.1369")
audit <- audit_everything(patience = 20, collections = "APSR")
```

After either kind of run, disk holds the **full** portfolio: the run merges into `audit_jobs.csv`, then rebuilds JSON and RDS from every CSV row.

### Seed without live runs

[refresh_registry()] (default `audit = FALSE`) seeds gaps so the health bar is not empty before the first full audit. It does **not** execute replications: prior `audit_latest.rds`, bake timings / artifact mtimes fill missing rows; existing `source = "audit"` rows are kept.

A normal `audit_everything()` already upserts and refreshes derived files; you do not need a separate rebuild after a live audit.

### `last_success_at`

Each CSV row carries `last_success_at`:

1. On a **successful** audit for that job → set to this run’s finish time.
2. Otherwise → keep the prior CSV value when present.
3. If still empty → fall back to bake timing `recorded_at`, else the local artifact’s mtime.

Failed or timed-out passes therefore do not erase the last known success time.

## How to read audit outputs

| Source | What it represents |
|--------|--------------------|
| `audit_jobs.csv` | Full portfolio job history (canonical). Inspect or edit here. |
| `audit_summary.json` | Baked counts + `progress` buckets for the **Shiny health bar**. Prefer this for first paint / lightweight consumers. |
| `audit_latest.rds` | Full `audit_everything` object derived from the **entire** CSV (not just the last call’s subset). |
| In-memory return of `audit_everything()` | **This run only** — if you passed `dois =` / `collections =`, `$results` and `$summary` cover that subset. Do not treat the return value as the portfolio when writing vignette snapshots or reports. |

For vignette or package snapshots, copy the registry RDS (full portfolio), not `saveRDS(audit)` from a filtered call:

```{r, eval = FALSE}
file.copy(
  file.path(getOption("replicateEverything.registry_root"), "audit_latest.rds"),
  "inst/vignette-data/audit_latest.rds",
  overwrite = TRUE
)
```

## Quarto report

The Quarto document in the registry repo still works the same way. Render it (or `scripts/run_audit.R`) to run a live audit and write the HTML report. After write, disk always holds the **full** portfolio CSV + derived JSON/RDS, even when params restrict `dois` / `collections`.

```{r, eval = FALSE}
# Quarto report lives in the registry repo (sibling registry/ in a monorepo)
quarto::quarto_render(audit_everything_qmd(), execute_params = list(patience = 20))
quarto::quarto_render(
  audit_everything_qmd(),
  execute_params = list(patience = 20, collections = "APSR")
)

# Or: live R audit, then HTML from saved RDS (no second live run)
# Rscript scripts/run_audit.R
# quarto render audit_everything.qmd -P refresh:false
```

## Console progress tags (0.7.39)

With `verbose = TRUE` (default), each finished job prints a short status tag aligned with the health-bar buckets:

| Tag | Meaning |
|-----|---------|
| `ok` | Runnable job succeeded (including substantive checks when defined) |
| `timeout` | Hit the `patience` audit cap |
| `substantive_fail` | Replication ran but published-value check failed |
| `missing_engine` | Engine / dependency gap (e.g. Stata not found) |
| `other` | Other failure or skip bucket |

Example line: `  - Table 1 (table, r) [ok]`.

## Commit checklist (registry repo)

After an audit (or seed + refresh) that you want on GitHub / Shiny:

1. `audit_jobs.csv` — always commit with the derived files
2. `audit_summary.json` — health bar
3. `audit_latest.rds` — full snapshot
4. `audit_everything.html` (+ `audit_everything_files/` if present) — only if you are publishing the Quarto report

Push the **registry** repository so remote consumers see the new summary. Optional: copy `audit_latest.rds` into `replicateEverything/inst/vignette-data/` and rebuild pkgdown so this article’s tables stay current.

## Latest audit results

The table below is built from a saved audit snapshot shipped with the package (`inst/vignette-data/audit_latest.rds`). Set `REPLICATE_AUDIT_LIVE=true` (and point at a local monorepo) only when you intend to refresh that snapshot.

```{r load-audit}
audit <- if (run_live) {
  tryCatch({
    monorepo <- Sys.getenv("REPLICATE_MONOREPO", unset = "")
    if (!nzchar(monorepo)) {
      parent <- normalizePath(
        file.path(find.package("replicateEverything"), "..", ".."),
        mustWork = FALSE
      )
      if (file.exists(file.path(parent, "registry", "index.csv"))) {
        monorepo <- parent
      }
    }
    if (nzchar(monorepo)) {
      options(
        replicateEverything.registry_root = file.path(monorepo, "registry"),
        replicateEverything.study_folders_root = monorepo,
        replicateEverything.use_sibling_packages = TRUE
      )
    }
    audit_everything(patience = 20, verbose = FALSE)
  }, error = function(e) {
    message("Live audit skipped: ", conditionMessage(e))
    NULL
  })
} else {
  NULL
}

if (is.null(audit) && nzchar(audit_rds) && file.exists(audit_rds)) {
  audit <- readRDS(audit_rds)
}
if (is.null(audit)) {
  stop("No audit results available.")
}
sm <- audit$summary
results <- audit$results
```

### Summary

| Metric | Value |
|:--|--:|
| Patience (seconds per object) | `r audit$patience` |
| Studies audited | `r sm$studies` |
| Replication runs | `r sm$runs` |
| Successful | `r sm$success` |
| Failed | `r sm$failed` |
| Timed out | `r sm$timed_out` |
| Skipped | `r sm$skipped %||% 0L` |
| Audit started | `r format(audit$started_at, "%Y-%m-%d %H:%M")` |
| Audit finished | `r format(audit$finished_at, "%Y-%m-%d %H:%M")` |

```{r pass-rate}
attempted <- sm$runs - (sm$skipped %||% 0L)
if (attempted > 0) {
  pct <- round(100 * sm$success / attempted, 1)
  cat(sprintf("**Pass rate (excluding skipped):** %s%%\n", pct))
}
```

### Results by study

```{r by-study, results = "asis"}
if (!"skipped" %in% names(results)) {
  results$skipped <- FALSE
}
studies <- unique(results$title)
for (study in studies) {
  cat("\n\n#### ", study, "\n\n", sep = "")
  sub <- results[results$title == study, , drop = FALSE]
  sub$status <- replicateEverything:::audit_result_status(
    sub$success, sub$timed_out, sub$skipped
  )
  sub$seconds <- ifelse(is.na(sub$seconds), NA, round(sub$seconds, 2))
  show <- sub[, c(
    "object_label", "object", "engine", "status", "seconds", "error_snippet"
  )]
  names(show) <- c("Object", "ID", "Engine", "Status", "Seconds", "Error")
  print(knitr::kable(show, row.names = FALSE))
}
```

### Failures (concise)

```{r failures}
if (!"skipped" %in% names(results)) {
  results$skipped <- FALSE
}
fails <- results[results$success %in% FALSE & !results$skipped %in% TRUE, , drop = FALSE]
if (nrow(fails) == 0) {
  cat("All recorded runnable jobs succeeded (or were skipped).\n")
} else {
  fails$seconds <- ifelse(is.na(fails$seconds), NA, round(fails$seconds, 2))
  fails$status <- replicateEverything:::audit_result_status(
    fails$success, fails$timed_out, fails$skipped
  )
  show <- fails[, c(
    "title", "object_label", "object", "engine",
    "status", "seconds", "error_snippet"
  )]
  names(show) <- c(
    "Study", "Object", "ID", "Engine", "Status", "Seconds", "Error"
  )
  knitr::kable(show, row.names = FALSE)
}
```

### Skipped (incomplete / unavailable)

```{r skips}
if (!"skipped" %in% names(results)) {
  results$skipped <- FALSE
}
skips <- results[results$skipped %in% TRUE, , drop = FALSE]
if (nrow(skips) == 0) {
  cat("No steps were skipped as incomplete / unavailable.\n")
} else {
  skips$status <- "Skipped"
  show <- skips[, c(
    "title", "object_label", "object", "engine", "status", "error_snippet"
  )]
  names(show) <- c(
    "Study", "Object", "ID", "Engine", "Status", "Reason"
  )
  knitr::kable(show, row.names = FALSE)
}
```

## Interpreting failures

Common reasons a run fails or times out:

- **Missing study package or folder** — install the study repo locally or set `replicateEverything.study_folders_root` to your monorepo root.
- **Stata not installed** — Stata-backed entries fail until Stata is found; see the *Stata replications* vignette.
- **Network / data** — folder-backed studies may need data files downloaded on first run.
- **Patience / audit cap** — slow jobs may time out under the configured `patience` (e.g. 60s in the registry report). That is recorded as **Timed out** with `timeout_seconds` and an explicit audit-cap message, not as a silent failure.
- **Skipped steps** — Mathematica / proprietary / incomplete yaml steps appear as **Skipped**, not Failed or Timed out.

Re-run locally and refresh the package vignette snapshot:

```{r, eval = FALSE}
Sys.setenv(REPLICATE_AUDIT_LIVE = "true")
options(
  replicateEverything.registry_root = "/path/to/replicate_everything/registry",
  replicateEverything.study_folders_root = "/path/to/replicate_everything",
  replicateEverything.use_sibling_packages = TRUE
)
audit <- audit_everything(patience = 20)
# Prefer the registry portfolio RDS (derived from audit_jobs.csv), not a
# filtered in-memory return value:
file.copy(
  file.path(getOption("replicateEverything.registry_root"), "audit_latest.rds"),
  "inst/vignette-data/audit_latest.rds",
  overwrite = TRUE
)

# Full HTML report: quarto render audit_everything.qmd (in the registry repo)
```

After adding or updating registry stubs, maintainers typically run [refresh_registry()] so `index.csv`, the Shiny cache, and the audit summary stay in sync:

```{r, eval = FALSE}
refresh_registry("../registry")                      # light
refresh_registry("../registry", audit = TRUE, patience = 20)
audit_report("../registry")
```
