---
title: "Folder replication checklist"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Folder replication checklist}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Two ways to contribute

| Model | Registry folder | Materials |
|-------|-----------------|-----------|
| **Registry paper** | `papers/<folder>/` with `code/`, `data/`, `artifacts/` | All in the registry repo |
| **Folder-backed** | `papers/<folder>/replication.yml` stub only | Study repo: `code/`, `data/`, `artifacts/` |

This checklist covers **folder-backed** studies (simple Git repo, not an R package).

## Study repo layout

```
rep-<doi-slug>/
  replication.yml
  README.md
  code/                 # one script per table/figure
  data/                 # optional
  artifacts/            # display outputs + manifest.json
  tests/
    testthat.R
    testthat/
      test-<id>.R
```

## Required `replication.yml` fields

**Paper metadata**

- `paper.doi` — full DOI URL
- `paper.title`
- `repo` or `paper.study_repo` — GitHub slug (`org/repo`); inferred from folder name when it starts with `rep-`

**Each figure or table**

```yaml
replications:
  - id: fig_1
    type: figure
    code: code/fig_1.R
    format: format_fig_1          # when display differs from analysis
    artifact: artifacts/fig_1.png
  - id: tab_1
    type: table
    code: code/tab_1.R
    format: format_tab_1
    data: data/myfile.dta
    artifact: artifacts/tab_1.html
```

## Workflow from the study repo

### 1. Build display artifacts

Run from the study repository root (or pass its path):

```{r, eval = FALSE}
library(replicateEverything)

options(
  replicateEverything.registry_root = "../registry",
  replicateEverything.use_sibling_packages = TRUE
)

build_study_artifacts(
  location = ".",
  install_deps = TRUE
)
```

This runs every replication, saves PNG/HTML (or RDS) under `artifacts/`, and writes `artifacts/manifest.json`.

### 2. Add tests

Each `tests/testthat/test-<id>.R` should:

- call `replicateEverything::run_replication(doi, id)` (analysis object)
- call `run_replication(..., format = TRUE)` when `format:` is set in yaml
- compare formatted output to the committed artifact (HTML normalize or PNG md5)

Run:

```{r, eval = FALSE}
testthat::test_dir("tests/testthat")
```

### 3. Validate before opening a PR

```{r, eval = FALSE}
check_folder_replication(
  location = ".",
  full_replication = FALSE,
  registry_root = "../registry"
)

# optional: also run every table/figure live
check_folder_replication(".", full_replication = TRUE, registry_root = "../registry")
```

### 4. Register with the registry

After checks pass, write the registry stub and update `index.csv`:

```{r, eval = FALSE}
options(replicateEverything.registry_root = "../registry")

add_folder_paper(
  location = ".",
  full_replication = FALSE
)
```

Remove any leftover `code/`, `data/`, or `artifacts/` under `registry/papers/<folder>/` — the registry keeps the stub only.

## Reference implementations

- [rep-10.1017-S0003055403000534](https://github.com/replicate-anything/rep-10.1017-S0003055403000534) (Fearon & Laitin)
- [rep-10.1177-00491241211036161](https://github.com/replicate-anything/rep-10.1177-00491241211036161) (Bounding Causes)

## Package-backed studies

For studies maintained as R packages, see `vignette("package-replication-checklist", package = "replicateEverything")` and `check_package_replication()` / `add_paper()`.
