---
title: "Folder replication checklist"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Folder replication checklist}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

See `vignette("why-replicateEverything", package = "replicateEverything")` for motivation and a tour of the consumer API.

## Two models

| Model | Registry | Materials |
|-------|----------|-----------|
| **Folder-backed** | `studies/<folder>.yml` stub only | Study repo: `code/`, `data/`, `outputs/` |
| **Package-backed** | `studies/<folder>.yml` stub only | Study R package on GitHub |

This checklist covers **folder-backed** studies (simple Git repo, not an R package).

## Study repo layout

```
rep-<doi-slug>/
  replication.yml
  README.md
  code/                 # one script per step
  data/                 # optional raw inputs
  outputs/              # step outputs + display files + manifest.json
  tests/
    testthat.R
    testthat/
      test-<id>.R
    substantive/          # published-value benchmarks (recommended)
      <step_id>.R
```

There is **no study-local `registry/` folder**. A maintainer with a local
registry checkout writes `studies/<folder>.yml` directly from this repo's
root `replication.yml` via `sync_study_to_registry()` (step 4) — nothing is
generated or committed here.

## Required `replication.yml` fields

**Paper metadata**

- `paper.doi` — full DOI URL
- `paper.title`
- `paper.article_url` — optional publisher landing page when `https://doi.org/...` does not resolve (common for older Cambridge Core / APSR entries)
- `repo` or `paper.study_repo` — GitHub slug (`org/repo`); inferred from folder name when it starts with `rep-`

**Maintainer and collections** (required for registry sync)

- `maintainer.name` and `maintainer.email` — person responsible for keeping the study repo and server dependencies current; shown as `[maintainer]` on the Studies tab
- `collections` — one or more tags for bibliography filtering (`APSR`, `PED`, `World Bank`, `IPI`, …); copied to registry `index.csv`
- `languages` — every engine used in steps (`r`, `stata`, `python`); precompiled into `index.csv` for the Studies tab

```yaml
maintainer:
  name: Jane Maintainer
  email: maintainer@example.org

collections:
  - APSR

languages:
  - r
  - stata
```

**Steps (required)**

Declare a single **`steps:`** DAG — yaml is the authority for execution order.
Legacy `prep:` / `replications:` blocks are a hard error; there is no
compatibility shim.

```yaml
steps:
  - id: prep_data
    type: transform
    parents: []
    inputs: [data/myfile.dta]
    outputs: [outputs/prep_data/analysis.rds]
    engine: r
    code: code/steps/prep_data.R

  - id: fig_1
    type: figure
    parents: [prep_data]
    data: outputs/prep_data/analysis.rds
    code: code/fig_1.R
    outputs: [outputs/fig_1.png]

  - id: fig_1_format
    type: format
    parent: fig_1
    code: code/format_fig_1.R
    outputs: [outputs/fig_1_display.png]
    # no label: — format children are sidebar-hidden wiring

  - id: tab_1
    type: table
    parents: [prep_data]
    code: code/tab_1.R
    outputs: [outputs/tab_1.html]
```

Format steps (`type: format`) run when `run_replication(..., format = TRUE)`.
Omit `label:` on format children (unused in Display).

## Workflow from the study repo

### 1. Build display artifacts

Run from the study repository root (or pass its path):

```{r, eval = FALSE}
library(replicateEverything)

options(
  replicateEverything.registry_root = "../registry",
  replicateEverything.use_sibling_packages = TRUE
)

build_study_outputs(
  location = ".",
  install_deps = TRUE
)
```

This runs every table and figure, saves display files under `outputs/`, and writes `outputs/manifest.json`.

Confirm outputs exist before handoff:

```{r, eval = FALSE}
validate_outputs(location = ".")
```

### 2. Add tests

Each `tests/testthat/test-<id>.R` should:

- call `replicateEverything::run_replication(doi, id)` (analysis object)
- call `run_replication(..., format = TRUE)` when a format child step exists
- compare formatted output to the committed file under `outputs/` (HTML normalize or PNG md5)

Run:

```{r, eval = FALSE}
testthat::test_dir("tests/testthat")
```

### 3. Validate (contributor)

```{r, eval = FALSE}
check_and_bake_study(
  location = ".",
  build_artifacts = FALSE   # already built above
)
```

This runs [check_replication()] against your study root `replication.yml`.
It **validates only** — it does not write a registry stub or any handoff
file into the study repo. Commit your study once checks pass; a registry
**maintainer** writes the stub directly from your yaml (step 4).

### 4. Sync to the central registry (maintainer)

Maintainers work from a local [registry](https://github.com/replicate-anything/registry) checkout:

```{r, eval = FALSE}
options(replicateEverything.registry_root = "../registry")

sync_study_to_registry(
  "../rep-10.1177-00491241211036161",
  registry_root = "../registry"
)
```

This builds a lightweight stub straight from the study repo's root
`replication.yml` (`maintainer`, `collections`, `languages`, and repo
pointers — never the full `steps:` list), writes it to
`registry/studies/<folder>.yml`, and runs [build_registry_index()].
`register_study(".", registry_root = "../registry")` runs
`check_and_bake_study()` then `sync_study_to_registry()` in one call.

After syncing several studies, refresh the whole registry and rerun the audit:

```{r, eval = FALSE}
refresh_registry("../registry", audit = TRUE, patience = 20)
```

### 5. Tests (recommended before syncing)

```{r, eval = FALSE}
testthat::test_dir("tests/testthat")
```

Run tests before `check_and_bake_study()` so the checklist reflects a merge-ready study.

### Substantive (published-value) checks

Smoke tests confirm that `run_replication()` returns an object. **Substantive checks** go further: they compare replicated estimates to **numbers in the published table** (coefficients, standard errors, sample sizes, etc.).

**Submitters:** add `tests/substantive/<step_id>.R` for each table or figure where you can quote benchmarks from the paper. Reference implementation: [Fearon & Laitin `tab_1`](https://github.com/replicate-anything/rep-10.1017-S0003055403000534/blob/main/tests/substantive/tab_1.R).

```r
# tests/substantive/tab_1.R — see Fearon & Laitin reference implementation
substantive_check_tab_1 <- function(object) {
  check_glm_table_benchmark(object, tab_1_prior_war_benchmark())
}
```

`check_glm_table_benchmark()` is injected when the package loads substantive check scripts (via [audit_everything()] or [check_replication()]). Study-specific benchmark values belong in the study repo, not the package.

Call the same function from `tests/testthat/test-<step_id>.R` so `testthat::test_dir()` exercises it locally.

**Maintainers:** [check_replication()] reports substantive coverage (`substantive_<step_id>` rows). Use `full_replication = TRUE` to run defined checks. [audit_everything()] runs them registry-wide by default (`substantive = TRUE`); failures appear as `[substantive]` in the audit summary.

### Maintainer shortcut

```{r, eval = FALSE}
options(replicateEverything.registry_root = "../registry")
sync_study_to_registry("../rep-10.1177-00491241211036161", audit = TRUE)
```

## Reference implementations

- [rep-10.1017-S0003055403000534](https://github.com/replicate-anything/rep-10.1017-S0003055403000534) (Fearon & Laitin)
- [rep-10.1177-00491241211036161](https://github.com/replicate-anything/rep-10.1177-00491241211036161) (Bounding Causes)

## Package-backed studies

For studies maintained as R packages, see `vignette("package-replication-checklist", package = "replicateEverything")` and `check_replication()`.
