---
title: "Maintainer setup"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Maintainer setup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

**replicateEverything** separates **readers** (browse, Run) from **maintainers** (onboard a study or server). Live Run and Shiny **probe** dependencies only — they never install R, Python, or Stata packages on the host. Maintainers install once during setup.

```{r setup}
library(replicateEverything)
```

## Check system compatibility (no installs)

Before running tables or figures, confirm the machine matches what `replication.yml` declares:

```{r}
check_study_compatibility("10.1017/S0003055426101749")
```

The returned list includes `ready` (logical), `install_needed`, and per-engine blocks under `dependencies` (`r`, `python`, `stata`).

In Shiny, use **Check system compatibility** above the tables list. When something is missing, a dialog shows the same maintainer commands documented here.

## Install dependencies for one study

Use **one function** for folder-backed and package-backed registry studies (R CRAN, Python pip, Stata `install_stata_deps.do`):

```{r}
install_dependencies("10.1017/S0003055426101749")
install_dependencies("10.1371/journal.pone.0278337")  # package-backed too
install_dependencies("path/to/study-repo")
```

This does **not** rebuild display artifacts — only dependencies. After setup, build artifacts with the kind-specific builder (paths differ by design):

```{r}
build_study_outputs("path/to/folder-study", install_deps = TRUE)
build_study_outputs("rep1371journalpone0278337", install_deps = TRUE)
```

Folder-backed studies write under `outputs/`; package-backed studies use `inst/report/outputs/` (or legacy `inst/report/artifacts/`).

## Install dependencies for every registry study

On a shared audit or Shiny server, run once:

```{r}
install_dependencies("everywhere")
```

Covers folder- and package-backed entries. Failures are collected per DOI; other studies continue.

## Point R at Python and Stata permanently

Session `options()` are not the right place for production servers. Set environment variables in **`~/.Renviron`** and restart R:

```text
PYTHON=C:/Users/you/AppData/Local/Python/pythoncore-3.14-64/python.exe
STATA=C:/Program Files/Stata17/StataMP-64.exe
```

On Linux or macOS, use the corresponding paths (for example `/usr/bin/python3` or `/Applications/Stata/StataMP.app/Contents/MacOS/stata-mp`).

In R you can open the file with `usethis::edit_r_environ()` if you use **usethis**.

Priority order:

| Engine | Permanent config | Fallback |
|--------|------------------|----------|
| Python | `PYTHON` or `RETICULATE_PYTHON` in `.Renviron` | Windows `py -0p` installs, then `PATH` |
| Stata | `STATA` or `REPLICATE_STATA_EXECUTABLE` in `.Renviron` | `options(replicateEverything.stata_executable)`, then common install paths |

## What happens on Run

When a reader calls `run_replication()` or Shiny **Run**, the package:

1. Checks declared dependencies (`assert_study_ready_for_replication()`).
2. Stops with [maintainer_dependency_hint()] if anything is missing.
3. Runs the replication **without** installing packages.

Maintainers see the same hint text in the console, in Shiny modals, and via `maintainer_dependency_hint()`:

```{r}
maintainer_dependency_hint("10.1017/S0003055426101749")
```

See also *Meet the functions*, the folder replication checklist, and `inst/ai/skills/include_study_in_registry.md` for contributor vs maintainer registry workflows.

## Register studies in the central registry (maintainer)

There is **no study-local registry handoff** — not `registry/` in a
folder-backed repo, not `inst/registry/` in a package. Contributors validate
with [check_and_bake_study()]; a maintainer with a local checkout of the
[registry repository](https://github.com/replicate-anything/registry) writes
the stub **directly from the study's own `replication.yml`** with
[sync_study_to_registry()] — nothing is copied from the study repo.

Point R at your monorepo checkout:

```{r}
options(replicateEverything.registry_root = "../registry")
```

**Sync one study** — build `registry/studies/<folder>.yml` from the study's
root yaml and rebuild `index.csv`:

```{r}
sync_study_to_registry(
  "../rep-10.1177-00491241211036161",
  registry_root = "../registry"
)
```

Optional per-study audit after sync:

```{r}
sync_study_to_registry(
  "../rep-10.1177-00491241211036161",
  registry_root = "../registry",
  audit = TRUE,
  patience = 20
)
```

**Refresh the whole registry** after a batch of syncs — recompile `index.csv` from all stubs and rerun [audit_everything()]:

```{r}
refresh_registry("../registry", audit = TRUE, patience = 20)
```

**One call, check + sync:** [register_study()] runs `check_and_bake_study()`
then `sync_study_to_registry()` for both folder- and package-backed studies:

```{r}
register_study(
  "../rep-10.1177-00491241211036161",
  registry_root = "../registry",
  audit = TRUE
)
```

## Check precomputed outputs

After building display files, confirm every declared table and figure exists on disk (Shiny **Display** mode). This does **not** run live replications.

```{r}
validate_outputs(location = "../rep-10.1177-00491241211036161")
validate_outputs("10.1177/00491241211036161", what = "everything")
```

Registry-wide check from a monorepo checkout:

```{r}
options(replicateEverything.registry_root = "../registry")
validate_outputs(doi = "everywhere", what = "everything")
```

Or from the registry repo: `Rscript scripts/validate_outputs.R`.

## Summary

| Task | Function |
|------|----------|
| Probe this machine | `check_study_compatibility(doi)` |
| Install one study (all kinds, all languages) | `install_dependencies(doi)` |
| Install entire registry | `install_dependencies("everywhere")` |
| Build study outputs | `build_study_outputs(location, install_deps = TRUE)` |
| Validate study | `check_replication(location)` |
| Check precomputed outputs | `validate_outputs(location)` or `validate_outputs(doi, what = "everything")` |
| Registry-wide output check | `validate_outputs(doi = "everywhere", what = "everything")` |
| Hint text for errors / UI | `maintainer_dependency_hint(doi)` |
| **Contributor:** validate (+ optional bake) | `check_and_bake_study(path)` |
| **Maintainer:** sync stub into registry | `sync_study_to_registry(path, registry_root = ...)` |
| **Maintainer:** validate then sync in one call | `register_study(path, registry_root = ...)` |
| **Maintainer:** rebuild index + audit all | `refresh_registry(registry_root, audit = TRUE)` |
| Rebuild index only | `build_registry_index(registry_root)` |

## Package website (pkgdown)

The site is served from **`docs/` on `main`** (GitHub Pages → `/docs`). CI does **not** build or deploy it.

```r
# from replicateEverything/
Rscript scripts/build_pkgdown.R
# then: git add docs/ && git commit && git push
```

Commit the **entire** `docs/` tree (`deps/`, `articles/`, `reference/`, …). Pushing only `index.html` breaks Bootstrap styling on GitHub.

**Windows + Dropbox:** if the repo lives under Dropbox, pkgdown may fail while writing `docs/articles/*.html` (e.g. `audit.html`) with `Invalid argument [1515]` / `Error closing file`. Pause Dropbox, exclude `docs/` from selective sync, or clone and build outside Dropbox. `scripts/build_pkgdown.R` deletes cached article HTML first when it detects a Dropbox path.
