---
title: "Maintainer setup"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Maintainer setup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

**replicateEverything** separates **readers** (browse, Run) from **maintainers** (onboard a study or server). Live Run and Shiny **probe** dependencies only — they never install R, Python, or Stata packages on the host. Maintainers install once during setup.

```{r setup}
library(replicateEverything)
```

## Check compatibility (no installs)

Before running tables or figures, confirm the machine matches what `replication.yml` declares:

```{r}
check_study_compatibility("10.1017/S0003055426101749")
```

The returned list includes `ready` (logical), `install_needed`, and per-engine blocks under `dependencies` (`r`, `python`, `stata`).

In Shiny, use **Check compatibility** above the tables list. When something is missing, a dialog shows the same maintainer commands documented here.

## Install dependencies for one study

Use a **single function** for all languages declared in the study repo (R CRAN, Python pip, Stata `install_stata_deps.do`):

```{r}
install_study_dependencies("10.1017/S0003055426101749")
# or a local clone:
install_study_dependencies("path/to/study-repo")
```

This does **not** rebuild `artifacts/` — it only satisfies dependencies. To build display artifacts after dependencies are in place:

```{r}
build_study_artifacts("path/to/study-repo", install_deps = TRUE)
```

Package-backed studies (registry entry with `paper.package`) use:

```{r}
build_package_artifacts("pkgname", install_deps = TRUE)
```

## Install dependencies for every registry study

On a shared audit or Shiny server, run once:

```{r}
install_registry_dependencies()
```

Failures are collected per DOI; other studies continue.

## Point R at Python and Stata permanently

Session `options()` are not the right place for production servers. Set environment variables in **`~/.Renviron`** and restart R:

```text
PYTHON=C:/Users/you/AppData/Local/Python/pythoncore-3.14-64/python.exe
STATA=C:/Program Files/Stata17/StataMP-64.exe
```

On Linux or macOS, use the corresponding paths (for example `/usr/bin/python3` or `/Applications/Stata/StataMP.app/Contents/MacOS/stata-mp`).

In R you can open the file with `usethis::edit_r_environ()` if you use **usethis**.

Priority order:

| Engine | Permanent config | Fallback |
|--------|------------------|----------|
| Python | `PYTHON` or `RETICULATE_PYTHON` in `.Renviron` | Windows `py -0p` installs, then `PATH` |
| Stata | `STATA` or `REPLICATE_STATA_EXECUTABLE` in `.Renviron` | `options(replicateEverything.stata_executable)`, then common install paths |

## What happens on Run

When a reader calls `run_replication()` or Shiny **Run**, the package:

1. Checks declared dependencies (`assert_study_ready_for_replication()`).
2. Stops with [maintainer_dependency_hint()] if anything is missing.
3. Runs the replication **without** installing packages.

Maintainers see the same hint text in the console, in Shiny modals, and via `maintainer_dependency_hint()`:

```{r}
maintainer_dependency_hint("10.1017/S0003055426101749")
```

## Summary

| Task | Function |
|------|----------|
| Probe this machine | `check_study_compatibility(doi)` |
| Install one study (all languages) | `install_study_dependencies(doi)` |
| Install entire registry | `install_registry_dependencies()` |
| Build artifacts after setup | `build_study_artifacts(path, install_deps = TRUE)` |
| Hint text for errors / UI | `maintainer_dependency_hint(doi)` |

See also *Meet the functions* and the folder replication checklist for study-repo conventions (`languages:`, `python_dependencies:`, `stata_deps_probe:`).
