---
title: "Meet the functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Meet the functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

**replicateEverything** has two sides. On the **consumer** side you browse the registry, run a table or figure, and inspect the code that produced it. On the **producer** side you set up a study repository, validate it, and register it so others can replicate your work.

This vignette walks through the main exported functions on each side. For full contributor checklists see the folder and package replication articles; for Stata-specific behaviour see *Stata replications*.

```{r setup}
library(replicateEverything)
```

## Using replicateEverything

These functions are for **readers and replicators**: you have a DOI (or a registry handle), you want to see what is available, run it, and read the scripts.

### Find a paper in the registry

Start with the index. `load_index()` returns registry `index.csv` as a data frame (DOI, title, journal, year, authors, study repo).

```{r}
idx <- load_index()
head(idx[, c("doi", "title", "year")])
```

Example output:

```
#>                              doi                                    title year
#> 1 10.1177/00491241211036161   Bounding Causes of Effects with Randomized... 2022
#> 2 10.1017/S0003055403000534   Ethnicity, Insurgency, and Civil War        2003
```

`search_papers()` keyword-searches titles and authors.

```{r}
search_papers("insurgency")
```

Example output:

```
#>                              doi                                  title
#> 1 10.1017/S0003055403000534 Ethnicity, Insurgency, and Civil War
```

For one paper, `list_replications()` lists every registered table and figure, including engine (`r` / `stata`) when declared.

```{r}
list_replications("10.1177/00491241211036161")
```

Example output (abbreviated):

```
#> [[1]]
#> $id
#> [1] "fig_1"
#> $type
#> [1] "figure"
```

`list_replication_groups()` returns **one entry per logical group**. When both R and Stata exist for the same table, the default is R.

```{r}
list_replication_groups("10.1257/aer.91.5.1369")
```

### Run a replication

`run_replication()` is the main entry point. Pass a DOI (or registry handle) and a replication id (`"fig_1"`, `"tab_1"`, and so on). By default you get the **analysis object** — a `ggplot`, model, or `data.frame`. Pass `format = TRUE` for display-ready HTML or a formatted plot.

```{r}
run_replication("10.1177/00491241211036161", "fig_1")

run_replication("10.1017/S0003055403000534", "tab_1", format = TRUE)

# Stata engine when both exist:
run_replication("10.1257/aer.91.5.1369", "tab_1", language = "stata", format = TRUE)
```

To reproduce an entire paper, set `what = "everything"`. The function runs every logical replication group (R by default where both engines exist).

```{r}
run_replication("10.1177/00491241211036161", "everything")
```

### View the replication code

`get_code()` returns the script behind a replication — R source or merged Stata `.do` files. Use it when you want to see *how* a result was produced, not only the output. The bundled Shiny app shows the same text on its Code tab.

```{r}
cat(get_code("10.1177/00491241211036161", "fig_1"), sep = "\n")
```

### Browse interactively in Shiny

The Shiny app is the consumer interface in a browser: pick a paper, switch between tables and figures, run replications, and copy the `run_replication()` call.

`run_shiny_app()` launches the bundled app from the installed package.

```{r}
run_shiny_app()
```

`save_local_shiny()` copies `app.R` and assets into a directory for Shiny Server (it never overwrites an existing `local.R`).

```{r}
save_local_shiny("/srv/shiny/replicate")
```

See `vignette("shiny-app", package = "replicateEverything")` for deployment details. A live demo runs at [shiny2.wzb.eu/ipi/replicate/](https://shiny2.wzb.eu/ipi/replicate/).

## Ready to contribute to replicateEverything?

These functions are for **authors and maintainers**: you have (or are building) a study repository with `replication.yml`, code, data, and artifacts, and you want to validate it and register it.

When you work inside a monorepo checkout (sibling `registry/` and `rep-*` study folders), point the package at local paths before calling discovery or run functions:

```{r}
options(
  replicateEverything.registry_root = "/path/to/replicate_everything/registry",
  replicateEverything.study_folders_root = "/path/to/replicate_everything",
  replicateEverything.use_sibling_packages = TRUE
)
```

Without these options the package reads the public registry from GitHub and clones study materials on demand.

### Maintainer setup (dependencies)

Live Run and Shiny **probe** dependencies only. Maintainers install once with:

```{r}
check_study_compatibility("10.1017/S0003055426101749")
install_study_dependencies("10.1017/S0003055426101749")  # folder or package DOI
install_registry_dependencies()
```

Build display artifacts separately: `build_study_artifacts()` (folder) or `build_package_artifacts()` (package). Inspect layout with `replication_kind()`. Full details: `vignette("maintainer-setup")`.

### Folder-backed studies

Study repos hold `replication.yml`, `code/`, `data/`, and `artifacts/`. See `vignette("folder-replication-checklist")`.

`build_study_artifacts()` runs every replication and writes PNG/HTML under `artifacts/` plus `manifest.json`.

```{r}
build_study_artifacts(".", install_deps = TRUE)
```

`check_folder_replication()` runs a transparent checklist: layout, yaml, artifacts, tests, and optional live runs.

```{r}
check_folder_replication(".", full_replication = FALSE)
```

`prepare_folder_paper()` builds artifacts (optional), runs checks, and writes `registry/replication.yml` and `registry/index.csv` in the study repo.

```{r}
prepare_folder_paper(".", build_artifacts = FALSE, registry_root = "../registry")
```

`sync_folder_paper()` copies prepared stub files into a local registry checkout.

```{r}
options(replicateEverything.registry_root = "../registry")
sync_folder_paper(".")
```

### Package-backed studies

Package-backed studies export `run_replication()`, `get_code()`, and related helpers from the study package itself. Validate with `check_package_replication()`.

```{r}
check_package_replication("../rep-10.1371_journal.pone.0278337")
check_package_replication("../rep-10.1371_journal.pone.0278337", full_replication = TRUE)
```

See `vignette("package-replication-checklist")` for layout and API requirements.

### Registry audit

`audit_everything()` attempts every table and figure in the registry (all engines), with a per-object time limit. Use it to check registry health after changes.

```{r}
audit <- audit_everything(patience = 20, dois = "10.1177/00491241211036161")
print(audit)
```

Example summary line:

```
#> Studies: 1 | Runs: 4 | OK: 4 | Failed: 0 | Timed out: 0
```

See `vignette("audit")` for the latest snapshot table shipped with the package.

## Quick reference

| Task | Function |
|------|----------|
| **Consumer** | |
| Browse registry | `load_index()`, `search_papers()` |
| What can I replicate? | `list_replications()`, `list_replication_groups()` |
| Run one result | `run_replication()` |
| Run whole paper | `run_replication(doi, "everything")` |
| View code | `get_code()` |
| Interactive browser | `run_shiny_app()`, `save_local_shiny()` |
| **Contributor** | |
| Check machine vs study yaml | `check_study_compatibility()` |
| Install deps (one study) | `install_study_dependencies()` |
| Install deps (all studies) | `install_registry_dependencies()` |
| Build folder artifacts | `build_study_artifacts()` |
| Validate folder study | `check_folder_replication()`, `prepare_folder_paper()`, `sync_folder_paper()` |
| Validate package study | `check_package_replication()` |
| Registry health check | `audit_everything()` |
