---
title: "Meet the functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Meet the functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

This vignette walks through the **main exported functions** in **replicateEverything**. For full contributor checklists see the folder and package replication articles; for Stata-specific behaviour see *Stata replications*.

## Local development setup

When you work inside a monorepo checkout (sibling `registry/` and `rep-*` study folders), point the package at local paths before calling discovery or run functions:

```{r}
options(
  replicateEverything.registry_root = "/path/to/replicate_everything/registry",
  replicateEverything.study_folders_root = "/path/to/replicate_everything",
  replicateEverything.use_sibling_packages = TRUE
)
```

Without these options the package reads the public registry from GitHub and clones study materials on demand.

```{r setup}
library(replicateEverything)
```

## Discovery

### `load_index()`

Returns the registry `index.csv` as a data frame (DOI, title, journal, year, authors, study repo).

```{r}
idx <- load_index()
head(idx[, c("doi", "title", "year")])
```

Example output:

```
#>                              doi                                    title year
#> 1 10.1177/00491241211036161   Bounding Causes of Effects with Randomized... 2022
#> 2 10.1017/S0003055403000534   Ethnicity, Insurgency, and Civil War        2003
```

### `search_papers()`

Keyword search over titles (and authors) in the index.

```{r}
search_papers("insurgency")
```

Example output:

```
#>                              doi                                  title
#> 1 10.1017/S0003055403000534 Ethnicity, Insurgency, and Civil War
```

### `list_replications()`

Lists every registered table and figure for one paper, including engine (`r` / `stata`) when declared.

```{r}
list_replications("10.1177/00491241211036161")
```

Example output (abbreviated):

```
#> [[1]]
#> $id
#> [1] "fig_1"
#> $type
#> [1] "figure"
```

### `list_replication_groups()`

Like `list_replications()`, but returns **one entry per logical group**. When both R and Stata exist for the same table, the default is R.

```{r}
list_replication_groups("10.1257/aer.91.5.1369")
```

### `get_code()`

Returns the replication script for the Code tab in Shiny (R source or merged Stata `.do` files).

```{r}
cat(get_code("10.1177/00491241211036161", "fig_1"), sep = "\n")
```

## Run replications

### `run_replication()`

Runs one table or figure. By default you get the **analysis object** (e.g. `ggplot`, model, or `data.frame`). Pass `format = TRUE` for display-ready HTML or a formatted plot.

```{r}
run_replication("10.1177/00491241211036161", "fig_1")

run_replication("10.1017/S0003055403000534", "tab_1", format = TRUE)

# Stata engine when both exist:
run_replication("10.1257/aer.91.5.1369", "tab_1", language = "stata", format = TRUE)
```

### `run_replication(..., what = "everything")`

Runs every logical replication group for a paper (R by default where bilingual).

```{r}
run_replication("10.1177/00491241211036161", "everything")
```

## Shiny demo

### `run_shiny_app()`

Launches the bundled Shiny app from the installed package.

```{r}
run_shiny_app()
```

### `save_local_shiny()`

Copies `app.R` and assets into a directory for Shiny Server (never overwrites an existing `local.R`).

```{r}
save_local_shiny("/srv/shiny/replicate")
```

See `vignette("shiny-app", package = "replicateEverything")` for deployment details.

## Contribute: folder-backed studies

Study repos hold `replication.yml`, `code/`, `data/`, and `artifacts/`. See `vignette("folder-replication-checklist")`.

### `build_study_artifacts()`

Runs every replication and writes PNG/HTML under `artifacts/` plus `manifest.json`.

```{r}
build_study_artifacts(".", install_deps = TRUE)
```

### `check_folder_replication()`

Transparent checklist: layout, yaml, artifacts, tests, optional live runs.

```{r}
check_folder_replication(".", full_replication = FALSE)
```

### `prepare_folder_paper()`

Builds artifacts (optional), runs checks, and writes `registry/replication.yml` and `registry/index.csv` in the study repo.

```{r}
prepare_folder_paper(".", build_artifacts = FALSE, registry_root = "../registry")
```

### `sync_folder_paper()`

Copies prepared stub files into a local registry checkout.

```{r}
options(replicateEverything.registry_root = "../registry")
sync_folder_paper(".")
```

## Contribute: package-backed studies

Package-backed studies export `run_replication()`, `get_code()`, and related helpers from the study package. Validate with:

```{r}
check_package_replication("../rep-10.1371_journal.pone.0278337")
check_package_replication("../rep-10.1371_journal.pone.0278337", full_replication = TRUE)
```

See `vignette("package-replication-checklist")` for layout and `check_package_replication()`.

## Registry audit

### `audit_everything()`

Attempts every table and figure in the registry (all engines), with a per-object time limit.

```{r}
audit <- audit_everything(patience = 20, dois = "10.1177/00491241211036161")
print(audit)
```

Example summary line:

```
#> Studies: 1 | Runs: 4 | OK: 4 | Failed: 0 | Timed out: 0
```

See `vignette("audit")` for the latest snapshot table shipped with the package.

## Quick reference

| Task | Function |
|------|----------|
| Browse registry | `load_index()`, `search_papers()` |
| What can I replicate? | `list_replications()`, `list_replication_groups()` |
| View code | `get_code()` |
| Run one result | `run_replication()` |
| Run whole paper | `run_replication(doi, "everything")` |
| Interactive browser | `run_shiny_app()`, `save_local_shiny()` |
| Folder study workflow | `build_study_artifacts()`, `check_folder_replication()`, `prepare_folder_paper()`, `sync_folder_paper()` |
| Package study workflow | `check_package_replication()` |
| Health check | `audit_everything()` |
