---
title: "Meet the functions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Meet the functions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

**replicateEverything** has two sides. On the **consumer** side you browse the registry, run a table or figure, and inspect the code that produced it. On the **producer** side you set up a study repository, validate it, and register it so others can replicate your work.

Start with the overview: `vignette("why-replicateEverything", package = "replicateEverything")`.

This vignette walks through the main exported functions on each side. For full contributor checklists see the folder and package replication articles; for Stata-specific behaviour see *Stata replications*.

```{r setup}
library(replicateEverything)
```

## Using replicateEverything

These functions are for **readers and replicators**: you have a DOI (or a registry handle), you want to see what is available, run it, and read the scripts.

### Find a paper in the registry

Start with the index. `load_index()` returns registry `index.csv` as a data frame (DOI, title, journal, year, authors, study repo).

```{r}
idx <- load_index()
head(idx[, c("doi", "title", "year")])
```

Example output:

```
#>                              doi                                    title year
#> 1 10.1177/00491241211036161   Bounding Causes of Effects with Randomized... 2022
#> 2 10.1017/S0003055403000534   Ethnicity, Insurgency, and Civil War        2003
```

`search_papers()` keyword-searches titles and authors.

```{r}
search_papers("insurgency")
```

Example output:

```
#>                              doi                                  title
#> 1 10.1017/S0003055403000534 Ethnicity, Insurgency, and Civil War
```

For one paper, `list_replications()` lists every registered table and figure, including engine (`r` / `stata`) when declared.

```{r}
list_replications("10.1177/00491241211036161")
```

Example output (abbreviated):

```
#> [[1]]
#> $id
#> [1] "fig_1"
#> $type
#> [1] "figure"
```

`list_replications(..., grouped = TRUE)` returns **one entry per logical group**. When both R and Stata exist for the same table, the default is R.

```{r}
list_replications("10.1257/aer.91.5.1369", grouped = TRUE)
list_replications("10.1257/aer.91.5.1369", grouped = TRUE, language = "stata")
```

Pipeline steps (transforms) are listed separately:

```{r}
list_replications("10.1017/s0003055426101749", include = "pipeline")
```

### Run a replication

`run_replication()` is the main entry point. Pass a DOI (or registry handle) and a replication id (`"fig_1"`, `"tab_1"`, and so on). By default you get the **analysis object** — a `ggplot`, model, or `data.frame`. Pass `format = TRUE` for display-ready HTML or a formatted plot.

```{r}
run_replication("10.1177/00491241211036161", "fig_1")

run_replication("10.1017/S0003055403000534", "tab_1", format = TRUE)

# Stata engine when both exist:
run_replication("10.1257/aer.91.5.1369", "tab_1", language = "stata", format = TRUE)
```

To reproduce an entire paper, set `what = "everything"`. The function runs every logical replication group (R by default where both engines exist).

```{r}
run_replication("10.1177/00491241211036161", "everything")
```

### View the replication code

`get_code()` returns the script behind a replication — R source or merged Stata `.do` files. Use it when you want to see *how* a result was produced, not only the output. The bundled Shiny app shows the same text on its Code tab.

```{r}
cat(get_code("10.1177/00491241211036161", "fig_1"), sep = "\n")
```

### Browse interactively in Shiny

The Shiny app is the consumer interface in a browser: pick a paper, switch between tables and figures, run replications, and copy the `run_replication()` call.

`run_shiny_app()` launches the bundled app from the installed package.

```{r}
run_shiny_app()
```

`save_local_shiny()` copies `app.R` and assets into a directory for Shiny Server (it never overwrites an existing `local.R`).

```{r}
save_local_shiny("/srv/shiny/replicate")
```

See `vignette("shiny-app", package = "replicateEverything")` for deployment details. A live demo runs at [shiny2.wzb.eu/ipi/replicate/](https://shiny2.wzb.eu/ipi/replicate/).

## Ready to contribute to replicateEverything?

These functions split **contributors** (authors preparing a study repo) from **maintainers** (registry operators syncing stubs and auditing the fleet).

When you work inside a monorepo checkout (sibling `registry/` and `rep-*` study folders), point the package at local paths before calling discovery or run functions:

```{r}
options(
  replicateEverything.registry_root = "/path/to/replicate_everything/registry",
  replicateEverything.study_folders_root = "/path/to/replicate_everything",
  replicateEverything.use_sibling_packages = TRUE
)
```

Without these options the package reads the public registry from GitHub and clones study materials on demand.

### Maintainer setup (dependencies)

Live Run and Shiny **probe** dependencies only. Maintainers install once with:

```{r}
check_study_compatibility("10.1017/S0003055426101749")
install_study_dependencies("10.1017/S0003055426101749")  # folder or package DOI
install_registry_dependencies()
```

Build display outputs with [build_study_outputs()]. Full details: `vignette("maintainer-setup")`.

### Folder-backed studies (contributor)

Study repos hold `replication.yml`, `code/`, `data/`, and `outputs/`. See `vignette("folder-replication-checklist")`.

[build_study_outputs()] runs every replication and writes display files under `outputs/` plus `outputs/manifest.json`.

```{r}
build_study_outputs(".", install_deps = TRUE)
```

[check_replication()] runs a transparent checklist: layout, yaml, outputs, tests, and optional live runs.

```{r}
check_replication(".", full_replication = FALSE)
```

`prepare_study_for_registry()` validates and writes `registry/replication.yml` + `registry/index.csv` **in the study repo** (contributor handoff).

```{r}
prepare_study_for_registry(".", build_artifacts = FALSE)
```

### Registry sync (maintainer)

Maintainers copy handoff files into the central registry checkout:

```{r}
options(replicateEverything.registry_root = "../registry")
sync_study_to_registry("../rep-10.1177-00491241211036161")
refresh_registry("../registry", audit = TRUE)
```

See `vignette("maintainer-setup")` for the full maintainer workflow.

### Check precomputed outputs

[validate_outputs()] checks that declared table and figure files exist on disk (for Shiny **Display**). It does not run live replications.

```{r}
validate_outputs(location = "../rep-10.1177-00491241211036161")
validate_outputs("10.1177/00491241211036161", what = "everything")
options(replicateEverything.registry_root = "../registry")
validate_outputs(doi = "everywhere", what = "everything")
```

### Package-backed studies (contributor)

Package-backed studies export `run_replication()`, `get_code()`, and related helpers from the study package itself. Validate with [check_replication()], then `prepare_study_for_registry()` for handoff files under `inst/registry/`.

```{r}
check_replication("../rep-10.1371_journal.pone.0278337")
check_replication("../rep-10.1371_journal.pone.0278337", full_replication = TRUE)
```

See `vignette("package-replication-checklist")` for layout and API requirements.

### Registry audit

`audit_everything()` attempts every table and figure in the registry (all engines), with a per-object time limit. Use it to check registry health after changes. Restrict with `dois =` or `collections =` (e.g. `"APSR"`).

```{r}
audit <- audit_everything(patience = 20, dois = "10.1177/00491241211036161")
# audit <- audit_everything(patience = 20, collections = "APSR")
print(audit)
```

Example summary line:

```
#> Studies: 1 | Runs: 4 | OK: 4 | Failed: 0 | Timed out: 0
```

See `vignette("audit")` for the latest snapshot table shipped with the package.

## Quick reference

| Task | Function |
|------|----------|
| **Consumer** | |
| Browse registry | `load_index()`, `search_papers()` |
| What can I replicate? | `list_replications()`, `list_replications(..., grouped = TRUE)` |
| Run one result | `run_replication()` |
| Run whole paper | `run_replication(doi, "everything")` |
| View code | `get_code()` |
| Interactive browser | `run_shiny_app()`, `save_local_shiny()` |
| **Contributor** | |
| Check machine vs study yaml | `check_study_compatibility()` |
| Install deps (one study) | `install_study_dependencies()` |
| Install deps (all studies) | `install_registry_dependencies()` |
| Build study outputs | `build_study_outputs()` |
| Validate study | `check_replication()` |
| Check precomputed outputs | `validate_outputs()` |
| Prepare registry handoff | `prepare_study_for_registry()` |
| **Maintainer** | |
| Sync study into registry | `sync_study_to_registry()` |
| Rebuild index + audit all | `refresh_registry()` |
| Rebuild index only | `build_registry_index()` |
| Registry output check | `validate_outputs(doi = "everywhere", what = "everything")` |
| Registry health check | `audit_everything()` |
