---
title: "Package replication checklist"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Package replication checklist}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Two models

| Model | Registry | Materials |
|-------|----------|-----------|
| **Folder-backed** | `studies/<folder>.yml` stub only | Study repo: `code/`, `data/`, `outputs/` |
| **Package-backed** | `studies/<folder>.yml` stub only | Study R package on GitHub |

This checklist covers **package-backed** studies (recommended for multi-table papers).

## Package layout

```
rep_<doi-slug>/
  DESCRIPTION
  replication.yml              # full metadata + replications list
  inst/replication.yml         # copy for installed package
  R/
    make_figure_*.R             # analysis
    make_table_*.R
    format_*.R                  # display formatting
    replication_api.R           # orchestration API (below)
  inst/replication_code/        # synced copies for Code tab / get_code()
  inst/registry/                 # handoff files (prepare_study_for_registry)
    replication.yml
    index.csv
  inst/report/artifacts/        # baked Display outputs (build_report())
  data/                         # analysis datasets (LazyData)
```

## Required `replication.yml` fields

**Paper metadata**

- `paper.doi` — full DOI URL
- `paper.title`
- `paper.package` — must match `DESCRIPTION` `Package:` field
- `paper.package_repo` or top-level `repo` — GitHub slug (`org/repo`)
- `paper.package_ref` — branch/tag (default `main`)
- `paper.package_folder` — optional; sibling folder name for monorepo dev

**Maintainer and collections** (required for registry sync)

- `maintainer.name` and `maintainer.email` — contact shown as `[maintainer]` on the Studies tab
- `collections` — tags for bibliography filtering (`APSR`, `PED`, `World Bank`, `IPI`, …)
- `languages` — engines used by the package (usually `r`)

```yaml
maintainer:
  name: Jane Maintainer
  email: maintainer@example.org

collections:
  - IPI

languages:
  - r
```

**Each figure or table**

```yaml
replications:
  - id: fig_1
    type: figure
    make: make_figure_1
    format: format_figure_1
    data: my_dataset
  - id: tab_1
    type: table
    make: make_table_1
    format: format_table_1
    data: my_dataset
```

Rules:

- `id` — short slug used by Shiny and `run_replication()`
- `type` — `figure` or `table`
- `make` — exported function returning analysis output (`ggplot` or `data.frame` / `kableExtra`)
- `format` — exported function returning display-ready output (`ggplot` or HTML string)
- `data` — package dataset name(s), not file paths

## Required exported API

Your package must export:

| Function | Purpose |
|----------|---------|
| `list_replications()` | Entries from `replication.yml` |
| `replication_meta()` | Parsed yaml |
| `run_replication(id)` | `make_*` then `format_*` |
| `load_artifact(id)` | Display HTML or PNG path |
| `artifact_file(id)` | Path to baked artifact file |
| `get_code(id)` | Source for Code tab |
| `build_report()` | Write `inst/report/artifacts/` |

## Artifacts (Display tab)

Run once before registering:

```r
build_report()
```

This writes:

- `inst/report/artifacts/<id>.png` for each figure
- `inst/report/artifacts/<id>.html` for each table (must contain a `<table>`)

The registry stub does **not** store artifacts.

## Prepare and register

### Contributor: validate and write handoff files

```{r, eval = FALSE}
library(replicateEverything)

prepare_study_for_registry(
  "../rep-10.1371_journal.pone.0278337",
  build_artifacts = TRUE
)
```

Writes `inst/registry/replication.yml` and `inst/registry/index.csv` in the package source tree. Commit these with your study PR.

### Maintainer: sync into the central registry

```{r, eval = FALSE}
options(replicateEverything.registry_root = "../registry")

sync_study_to_registry(
  "../rep-10.1371_journal.pone.0278337",
  registry_root = "../registry",
  audit = TRUE
)

# After several syncs:
refresh_registry("../registry", audit = TRUE)
```

### Checks only

```{r, eval = FALSE}
check_replication(
  "../rep-10.1371_journal.pone.0278337",
  full_replication = FALSE
)

check_replication(
  "../rep-10.1371_journal.pone.0278337",
  full_replication = TRUE
)
```

## Substantive (published-value) checks

Package-backed studies use the same `tests/substantive/<step_id>.R` convention as folder-backed repos. Define `substantive_check_<step_id>(object)` and call it from `tests/testthat/`. [check_replication()] reports coverage; `full_replication = TRUE` runs defined checks. [audit_everything()] includes them in the registry audit (`substantive = TRUE`, default).

See Fearon & Laitin Table 1 in [rep-10.1017-S0003055403000534](https://github.com/replicate-anything/rep-10.1017-S0003055403000534) and `?check_glm_table_benchmark`.

## Reference implementation

See [rep-10.1371_journal.pone.0278337](https://github.com/replicate-anything/rep-10.1371_journal.pone.0278337) (vaccine solidarity paper).
