---
title: "Reanalysis and extension studies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reanalysis and extension studies}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

A **reanalysis study** reuses upstream data preparation from a published replication while replacing one or more analysis steps. The replicateEverything package supports this with `paper.extends` and step-level `inherit:` declarations.

The worked example pairs:

- **Base study:** [Fearon & Laitin (2003)](https://doi.org/10.1017/S0003055403000534) — `rep-10.1017-S0003055403000534`
- **Reanalysis:** [Simple replication of F&L data / repo](https://github.com/replicate-anything/rep-10.1017-S0003055403000534--alt-1) — same prepared data, `lm_robust` instead of `glm`

## Base study pipeline

The base study declares a transform step and downstream tables:

```
raw data/repdata.dta  →  prep_data  →  tab_1 (R) / tab_1_stata (Stata)
```

`prep_data` renames `lpopl1` to `lpopl` and recodes onset indicators. Both analysis engines read `outputs/prep_data/repdata.rds` or `.dta`.

## Extension study layout

The reanalysis repository holds only new material. When the display format matches the base study, inherit the format child step; when it differs (as here), inherit with a \code{code:} override pointing at local R scripts:

```yaml
steps:
  - inherit: prep_data

  - id: tab_1
    type: table
    parents: [prep_data]
    data: outputs/prep_data/repdata.rds
    code: code/tab_1.R
    format: format_tab_1

  - inherit: tab_1_format
    code: code/tab_1.R
```

In the base Fearon \& Laitin study, Stata table steps read \code{data/repdata.dta} directly; R steps use the shared \code{prep_data} output.

## Execution semantics

| `given` | Behaviour |
|---------|-----------|
| `"parents"` | Requires `prep_data` outputs in the **base** repo (`outputs/prep_data/`) |
| `"nothing"` | Runs inherited `prep_data` in the base checkout, then the extension analysis locally |

Inherited steps execute in the base study root; extension steps run in the extension root but may read base `outputs/`.

## Running locally

From a monorepo checkout with both study folders as siblings:

```{r, eval = FALSE}
devtools::load_all("replicateEverything")
configure_local_monorepo()

# Base study (once)
run_replication("10.1017/S0003055403000534", "prep_data", given = "nothing")

# Reanalysis (uses base prep_data outputs)
run_replication(
  "rep-10.1017-S0003055403000534--alt-1",
  "tab_1",
  given = "parents"
)
```

Or pass the extension study path directly when no article DOI exists yet:

```{r, eval = FALSE}
run_replication("rep-10.1017-S0003055403000534--alt-1", "tab_1")
```

## Registration

Extension studies without a DOI use `paper.study_handle` in `replication.yml` and a registry stub keyed by that handle. Link to the original study in `paper.related` or the abstract. See `registry/studies/rep-10.1017-S0003055403000534--alt-1.yml` for the stub template.

Validate before opening a PR:

```{r, eval = FALSE}
check_folder_replication("rep-10.1017-S0003055403000534--alt-1")
```

## Further reading

- Step inheritance design notes: `system.file("docs/step-inheritance.md", package = "replicateEverything")`
- Folder replication checklist: `vignette("folder-replication-checklist", package = "replicateEverything")`
