---
title: "Replication Examples Using Code"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Replication Examples Using Code}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

```{r setup}
library(replicateEverything)
```

## Quick start

```{r}
# Look up bibliographic metadata
get_doi_metadata("10.1177/00491241211036161")

# Search the registry by title keyword
search_papers("causes")

# See what can be replicated for a paper
list_replications("10.1177/00491241211036161")

# Run one figure or table
run_replication("10.1177/00491241211036161", "fig_1")

# Reproduce every registered result
replicate_paper("10.1177/00491241211036161")
```

## How it works

The [registry](https://github.com/replicate-anything/registry) indexes studies with lightweight stub files in `papers/<folder>.yml` and `index.csv`. **Replication materials live in separate study repositories**, not inside the registry.

`replicateEverything` reads the stub, fetches the full `replication.yml` from the study repo or package, loads data, sources analysis scripts, and returns typed result objects.

```
Registry (index only)              Study repository
  papers/<folder>.yml  ───────►    replication.yml
  index.csv                        data/
                                   code/
                                   artifacts/   (folder-backed)
              ↓
      replicateEverything
              ↓
      figures & tables in your R session
```

### Two study layouts

**Folder-backed** studies use a simple Git repository:

```
rep-<doi-slug>/
  replication.yml
  data/
  code/
  artifacts/
  tests/testthat/
```

**Package-backed** studies use a standalone R package:

```
rep_<doi_slug>/
  DESCRIPTION
  R/
  data/
  replication.yml
  inst/report/artifacts/
```

In both cases the registry holds only a **stub** (`papers/<folder>.yml`) pointing at the study repo. See the folder and package replication checklists for contributor workflows.

### Shiny demo

A [live demo](https://shiny2.wzb.eu/ipi/replicate/) runs at WZB. The package also bundles the app in `inst/shiny/`:

```{r}
run_shiny_app()
save_local_shiny("/path/to/shiny/replicate")
```

See `vignette("shiny-app", package = "replicateEverything")` for deployment on Shiny Server.

## System architecture

```{r, fig_1_example, echo=FALSE, fig.width=6, fig.height=4, out.width="70%", fig.align="center", eval = FALSE}
suppressPackageStartupMessages(library(DiagrammeR))
DiagrammeR::grViz("
digraph {
  Registry -> 'Study repositories'
  'Study repositories' -> replicateEverything
  replicateEverything -> Shiny
}
")
```

## Run a single replication

```{r}
run_replication("10.1177/00491241211036161", "fig_1")
```

## Replicate an entire paper

```{r}
replicate_paper("10.1177/00491241211036161")
```
