---
title: "Shiny demo app"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Shiny demo app}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

The **replicateEverything** package includes a Shiny demo in `inst/shiny/`
(`app.R` plus `www/` assets). You can run it directly from the installed
package, or copy it into a Shiny Server directory.

## Install

```{r}
# install.packages("remotes")
remotes::install_github("replicate-anything/replicateEverything")
```

Suggested packages for the app: `shiny`, `bslib`, and optionally `shinyWidgets`.

```{r}
install.packages(c("shiny", "bslib", "shinyWidgets"))
```

## Option 1: Run from the package

```{r}
library(replicateEverything)
run_shiny_app()
```

This launches `inst/shiny` inside the installed package. The app uses the
installed `replicateEverything` version and does not try to reinstall itself
from GitHub.

For local monorepo development (sibling `registry/` and study packages), copy
`inst/shiny/local.R.example` to `local.R` in your working directory before
calling `run_shiny_app()`, or set options manually:

```{r}
options(
  replicateEverything.registry_root = "/path/to/registry",
  replicate_shiny.auto_update_replicate_everything = FALSE
)
```

## Option 2: Copy for Shiny Server

Many servers expect a directory with `app.R` (for example `shiny/replicate/`).
After installing or updating the package, materialize the bundled app:

```{r}
library(replicateEverything)
save_local_shiny("/srv/shiny/replicate")
```

This writes:

- `app.R`
- `www/` (logo and favicons)
- `local.R.example` (template only)

**`local.R` is never overwritten**, so server-specific settings survive updates.

### Server update workflow

```{r}
remotes::install_github("replicate-anything/replicateEverything")
library(replicateEverything)
save_local_shiny("/srv/shiny/replicate")
```

Restart the app on your host if required.

### Server configuration

On a shared server, create `local.R` once (from `local.R.example`):

```{r}
options(
  replicateEverything.registry_root = "/path/to/registry",
  replicate_shiny.auto_update_replicate_everything = FALSE,
  replicate_shiny.auto_install_study_packages = FALSE
)
```

If you rely on the public GitHub registry, you do not need a local `registry/`
checkout; omit `replicateEverything.registry_root`.

## Former standalone repository

The app previously lived in a separate **replicate-shiny** repository beside
**replicateEverything**. That repository is deprecated; use the functions above.
