---
title: "Shiny demo app"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Shiny demo app}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

The **replicateEverything** package includes a Shiny demo in `inst/shiny/`
(`app.R` plus `www/` assets). A live instance runs at
<https://shiny2.wzb.eu/ipi/replicate/>. You can also run the app from an
installed package, or copy it into a Shiny Server directory.

## Install

```{r}
# install.packages("remotes")
remotes::install_github("replicate-anything/replicateEverything")
```

Suggested packages for the app: `shiny`, `bslib`, and optionally `shinyWidgets`.

```{r}
install.packages(c("shiny", "bslib", "shinyWidgets"))
```

## Option 1: Run from the package

```{r}
library(replicateEverything)
run_shiny_app()
```

This launches `inst/shiny` inside the installed package. The app uses the
installed `replicateEverything` version and does not try to reinstall itself
from GitHub.

For local monorepo development (sibling `registry/` and study packages), copy
`inst/shiny/local.R.example` to `local.R` in your working directory before
calling `run_shiny_app()`, or set options manually:

```{r}
options(
  replicateEverything.registry_root = "/path/to/registry",
  replicate_shiny.auto_update_replicate_everything = FALSE
)
```

## Option 2: Copy for Shiny Server

Many servers expect a directory with `app.R` (for example `shiny/replicate/`).
After installing or updating the package, materialize the bundled app:

```{r}
library(replicateEverything)
save_local_shiny("/srv/shiny/replicate")
```

This writes:

- `app.R`
- `www/` (logo and favicons)
- `local.R.example` (template only)
- `deploy-options.R` (display-only vs live run; always overwritten on deploy)
- `BUNDLE_SHA` (package build stamp)

**`local.R` is never overwritten**, so server-specific settings survive updates.

### Display-only vs Live Run

By default, deployed apps allow **Live Run** (same as local `run_shiny_app()`).
For a public demo that should only show precomputed artifacts, deploy in
display-only mode:

```{r}
save_local_shiny("/srv/shiny/replicate", live_run = FALSE)
```

This writes `deploy-options.R` with `options(replicate_shiny.live_run = FALSE)`.
The app hides Run buttons and shows a subtle banner. Use `live_run = TRUE`
(the default) when the server should execute replications on demand.

For local development, set `options(replicate_shiny.live_run = TRUE)` in
`local.R` if you copied a display-only deploy bundle but want Live Run while
developing (see `local.R.example`).

### Server update workflow

```{r}
remotes::install_github("replicate-anything/replicateEverything")
library(replicateEverything)
save_local_shiny("/srv/shiny/replicate")
```

Restart the app on your host if required.

### Display artifacts

Shiny **Display** mode serves precomputed files from each study's `outputs/` (folder-backed) or the study package's report outputs. Maintainers can verify those files exist with [validate_outputs()] — per study (`location =` or `doi` + `what = "everything"`) or registry-wide (`doi = "everywhere"`, `what = "everything"`). See `vignette("maintainer-setup")`.

### Server configuration

On a shared server, create `local.R` once (from `local.R.example`):

```{r}
options(
  replicateEverything.registry_root = "/path/to/registry",
  replicate_shiny.auto_update_replicate_everything = FALSE,
  replicate_shiny.auto_install_study_packages = FALSE
)
```

If you rely on the public GitHub registry, you do not need a local `registry/`
checkout; omit `replicateEverything.registry_root`.

## Code tab: inspect sourced files

On the **Code** tab, Stata runners are shown as authored (not inlined). Lines
such as `do "${maindir}/code/tables/mk_tab_1.do"` are clickable when the target
file exists under the study root. Breadcrumb navigation and **Back** let you
walk nested `do` / `source()` calls without stacking modals.

**Path globals:** folder-backed Stata studies set `global maindir` in
`code/helpers/init_study_paths.do` to the directory containing `replication.yml`
(walked up from the working directory at run time). The Shiny viewer uses the
same mapping via `default_stata_globals(study_root)` (`maindir`, `rawdir`,
`processed`, `result`). Live runs may override `result` with
`REPLICATE_STATA_RESULT`.

Implementation lives in `R/code_links.R` (`build_code_file_graph()`,
`render_code_html_with_links()`). A future `code_manifest:` block in
`replication.yml` may point at Dataverse-hosted scripts (similar to the data
manifest pattern).

**Not yet parsed:** unquoted `do` paths with embedded spaces, `` `local' `` /
compound double quotes, `source(file.path(...))`, and Python `exec`/`runpy`.

## Former standalone repository

The app previously lived in a separate **replicate-shiny** repository beside
**replicateEverything**. That repository is deprecated; use the functions above.
