---
title: "Stata replications"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Stata replications}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Overview

**replicateEverything is an R package.** It orchestrates replications from R: loading registry metadata, fetching study folders or packages, and running code. Stata-backed entries are **first-class**: the package calls Stata in batch mode, captures logs or images, and can format them for display in R or Shiny.

You can also register **bilingual** studies with paired R and Stata entries (same `group`, different `engine`). The [Acemoglu et al.](https://doi.org/10.1257/aer.91.5.1369) study in the registry is an example: each table has `tab_N` (R) and `tab_N_stata` (Stata).

## How Stata replications work in R

1. **`replication.yml`** marks an entry with `engine: stata` (or a `.do` file path).
2. **`run_replication()` / `render_replication()`** detect the engine and call **`run_stata_replication()`**.
3. R locates Stata via **`find_stata_executable()`**, writes a small runner, and invokes Stata batch mode (`StataMP-64.exe /e do ...` on Windows).
4. Output is returned as an R object of class **`stata_replication_result`** — a list with (at minimum) `output_path` (log or image), `study_root`, and `replication_id`.
5. If the entry defines `format: code/format_stata.R`, R reads the Stata log and builds an HTML table for Shiny / `format = TRUE`.

The returned object is **not** a Stata dataset inside R; it is a **handle to Stata output on disk** (usually a `.log` file). Use `replication_object()` to extract it.

## Finding Stata

```{r, eval = FALSE}
library(replicateEverything)

find_stata_executable()
```

If this returns `NULL`, R will stop with a clear error when you run a Stata replication:

> Replication tab_1_stata requires Stata. Install Stata or set options(replicateEverything.stata_executable = ...).

**Point R at your installation:**

```{r, eval = FALSE}
options(replicateEverything.stata_executable = "C:/Program Files/Stata18/StataMP-64.exe")
# macOS example:
# options(replicateEverything.stata_executable = "/Applications/Stata/StataMP.app/Contents/MacOS/stata-mp")
```

`find_stata_executable()` checks, in order:

1. `getOption("replicateEverything.stata_executable")`
2. `PATH` (`stata-mp`, `stata-se`, …)
3. Common install locations (Windows Program Files, `/Applications/Stata/…`, Linux `/usr/local/stata*`)

## Running from R

```{r, eval = FALSE}
configure_local_monorepo("/path/to/replicate_everything")

# Stata only
run_replication("10.1257/aer.91.5.1369", "tab_1_stata")

# Formatted HTML table (uses format_stata.R in the study repo)
run_replication("10.1257/aer.91.5.1369", "tab_1_stata", format = TRUE)

# Inspect the raw result object
result <- render_replication("10.1257/aer.91.5.1369", "tab_1_stata")
str(replication_object(result))
```

List engines available for a study:

```{r, eval = FALSE}
reps <- list_replications("10.1257/aer.91.5.1369")
table(vapply(reps, function(x) x$engine %||% "r", character(1)))
```

## Running the same code in Stata yourself

replicateEverything is for **orchestration from R**. To work in native Stata (debugging, extending analysis), use the study repository directly.

### 1. Clone the study repo

```         
git clone https://github.com/replicate-anything/rep-10.1257-aer.91.5.1369.git
cd rep-10.1257-aer.91.5.1369
```

Data files live under `data/` (e.g. `maketable1.dta`). They are referenced from `replication.yml` and must be present locally (download from the study’s `source_url` if not in git).

### 2. Table 1 — runner vs analysis

The registry uses two layers:

| File | Role |
|------|------|
| `code/tab_1.do` | **Runner** — sets paths, opens a log, calls the analysis |
| `code/maketable1.do` | **Analysis** — MIT original code (`summ` blocks for Table 1) |

From the **study repo root** in Stata:

```         
do "code/tab_1.do"
```

The runner sets `global maindir` to the repo root, `global datadir` to `data/`, and writes `artifacts/staging/tab_1_stata.log`. That matches what R runs via `run_stata_replication()`.

### 3. View code from R before opening Stata

```{r, eval = FALSE}
cat(get_code("10.1257/aer.91.5.1369", "tab_1_stata"), sep = "\n")
```

`get_code()` merges the runner with `stata_source` (`code/maketable1.do`) so you see the full script Shiny’s Code tab would show.

### 4. Minimal manual setup (without the runner)

If you already have the repo open in Stata:

```         
global maindir "C:/path/to/rep-10.1257-aer.91.5.1369"
global datadir "${maindir}/data"
do "${maindir}/code/maketable1.do"
```

Use forward slashes or Stata’s path conventions on your OS.

## Bilingual entries in `replication.yml`

Acemoglu Table 1 is declared twice:

```yaml
  - id: tab_1
    group: tab_1
    engine: r
    code: code/tab_1.R
    ...

  - id: tab_1_stata
    group: tab_1
    engine: stata
    code: code/tab_1.do
    stata_source: code/maketable1.do
    format: code/format_stata.R
    output: artifacts/staging/tab_1_stata.log
```

Shiny and `audit_everything()` treat `group` as one logical table and run **both** engines when listed.

## Example: Table 1 from R

```{r acemoglu-tab1, eval = FALSE}
library(replicateEverything)

# Optional: local monorepo + Stata path
configure_local_monorepo("c:/path/to/replicate_everything")
options(replicateEverything.stata_executable = "C:/Program Files/Stata18/StataMP-64.exe")

# R translation
run_replication("10.1257/aer.91.5.1369", "tab_1", format = TRUE)

# Original Stata (when Stata is installed)
run_replication("10.1257/aer.91.5.1369", "tab_1_stata", format = TRUE)
```

When Stata is unavailable, the R entry still runs; Stata entries fail gracefully with an install hint.

## Related reading

- [Folder replication checklist](folder-replication-checklist.html) — layout for folder-backed studies, including `engine: stata`
- [Registry audit](audit.html) — `audit_everything()` runs all R and Stata entries with a per-object time limit
