---
title: "Why replicateEverything?"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Why replicateEverything?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
live <- identical(tolower(Sys.getenv("REPLICATE_VIGNETTE_LIVE", unset = "")), "true")
```

<style>
.tldr-box {
  border: 1px solid #e0e0e0;
  border-radius: 10px;
  padding: 1.1em 1.3em;
  margin: 1.6em 0 2em 0;
  box-shadow: 0 6px 18px rgba(0, 0, 0, 0.10);
  background: #fbfbfb;
}
.tldr-box .tldr-label {
  font-weight: 600;
  font-size: 1.05em;
  margin-bottom: 0.6em;
  letter-spacing: 0.01em;
}
</style>

<div class="tldr-box">

<p class="tldr-label">TL;DR — just do this</p>

```{r tldr-code, eval=live, fig.height=4.5, fig.width=6, fig.align='left', message=FALSE, warning=FALSE}
library(replicateEverything)
options(replicateEverything.quiet_run = TRUE)
fig <- replicateEverything::run_replication(
  doi = "10.1017/s0003055426101749",
  what = "fig_4",
  format = TRUE
)
if (inherits(fig, "ggplot")) {
  print(fig)
} else {
  plot(1, 1, type = "n", axes = FALSE, xlab = "", ylab = "")
  text(1, 1, "Set REPLICATE_VIGNETTE_LIVE=true to render fig_4")
}
```

```{r tldr-code-static, eval=!live, echo=TRUE}
library(replicateEverything)
replicateEverything::run_replication(
  doi = "10.1017/s0003055426101749",
  what = "fig_4"
)
```

</div>

**replicateEverything** is a prototype system to:

- structure replication code as a connected pipeline
- fetch and run that code from a DOI (or registry handle)
- check that it still works — one study, or a whole journal at once
- link base replications to downstream re-analyses without copying files

You can use it from `R` or from the browser: [shiny2.wzb.eu/ipi/replicate/](https://shiny2.wzb.eu/ipi/replicate/).

## Why bother?

Computational replication has never been easier to *attempt*; it is still surprisingly hard to *repeat* someone else's analysis on demand. Archives are ad hoc, paths differ, dependencies drift, and every new reader retraces the same scavenger hunt. Nobody runs a standing check that the archive still runs today.

Are we fighting the lat battle. You might wonder: if an AI can rewrite the analysis from the paper, why keep an authoritative codebase? Our answe: Because the code *is* the record of analytic choices. When results disagree, you need a shared artifact to inspect, not a fresh vibe-coded reimplementation. That case gets stronger  as models get better at improvising.

## The registry (what is in there so far)

The [registry](https://github.com/replicate-anything/registry) indexes studies with lightweight stubs; code and data live in linked study repos. The index is still small — but you can browse what is there:

```{r index-head, eval=FALSE}
library(replicateEverything)
head(load_index()[, c("doi", "title", "year", "collections")])
```

Example output:

```
#>                              doi                                    title year collections
#> 1 10.1177/00491241211036161   Bounding Causes of Effects ...          2022 PED
#> 2 10.1017/S0003055403000534   Ethnicity, Insurgency, and Civil War    2003 APSR
#> 3 10.1017/s0003055426101749   ...                                     2026 APSR
```

## Run one result

Pass a DOI and a step id (`fig_1`, `tab_1`, …):

```{r basic-run, eval=FALSE}
library(replicateEverything)

run_replication(
  doi = "10.1017/S0003055403000534",
  what = "tab_1"
)

# Stata when both engines exist:
run_replication(
  doi = "10.1257/aer.91.5.1369",
  what = "tab_2",
  language = "stata"
)
```

`list_replications()` shows what you can run — now with a compact print method instead of dumping nested lists:

```{r list-reps, eval=FALSE}
list_replications("10.1017/S0003055403000534")
# Replications: Ethnicity, Insurgency, and Civil War [10.1017/s0003055403000534]
#        id   type engine                      label
#     tab_1  table      r                     Table 1
# tab_1_stata table  stata                     Table 1
```

Use `grouped = TRUE` when you want one row per logical table or figure (R by default when both exist).

Pull the script any time:

```{r get-code, eval=FALSE}
get_code("10.1017/S0003055403000534", "tab_1")
```

Run every non-format step in the study pipeline:

```{r everything, eval=FALSE}
run_replication("10.1017/S0003055403000534", "everything")
```

For `what = "everything"`, upstream prep runs automatically (`given` defaults to `"nothing"`). For a single table or figure, the default is `"parents"` — use existing intermediate outputs when they are already on disk.

Pass `format = TRUE` for display-ready HTML or formatted plots; the default returns raw analysis objects (models, `ggplot`s, and so on).

## Wrinkles worth knowing

### Multi-step pipelines (DAGs)

A serious archive is a **DAG**: raw data → transforms → tables and figures. `replicateEverything` reads that graph from `replication.yml` so you can start from prepared data or from scratch.

```{r dag, eval=FALSE}
describe_study_dag("10.1017/S0003055403000534")
```

See *Reanalysis and extension studies* for downstream repos that inherit upstream steps without duplicating code.

### Reanalysis without redundancy

Extension studies set `paper.extends` and `inherit:` steps. Inherited prep runs in the base repo; new analysis reads base `outputs/`. Details: `vignette("reanalysis-studies")`.

### Many languages, one entry point

Folder-backed studies may mix R, Stata, and Python. Pick the engine with `language =` when both exist. See `vignette("stata-replications")`.

### Collections

Registry rows carry `collections` tags (`APSR`, `PED`, `IPI`, `World Bank`, …) for filtering in the Shiny bibliography.

## Using AI to prep your archive

The fastest path from a messy delivery to a registry-ready study is often an agent with the package **skills** — markdown playbooks shipped in `inst/ai/skills/`:

```{r ai-skills, eval=FALSE}
ai_skills()
# folder-replication       — generic folder-backed study repo
# apsr-to-replicate-everything — APSR / Cambridge Dataverse flat deliveries
```

Point Cursor (or another agent) at the skill, give it your replication folder, and ask it to produce `replication.yml` with a `steps:` DAG, `outputs/` paths, tests, and a registry stub. The skills spell out layout, dependency probing, substantive tests, and `check_folder_replication()`.

Practical tips without an agent:

1. **Start from the published code**, not a rewrite — discover the true pipeline (what reads what) before you yaml it.
2. **One script per step**; declare inputs, outputs, and parents in `steps:`.
3. **Run** `check_folder_replication()` and `prepare_folder_paper()` before opening a registry PR.
4. **Add substantive tests** under `tests/substantive/` when you have published benchmarks (see Fearon & Laitin `tab_1`).

Skill files live at `system.file("ai/skills", package = "replicateEverything")` after install.

## Where to go next

| If you want to… | Read |
|-----------------|------|
| Tour every main function | `vignette("meet-the-functions")` |
| Run examples from code | `vignette("replication-example")` |
| Build or migrate a folder-backed study | `vignette("folder-replication-checklist")` |
| Onboard an APSR / Dataverse delivery | skill `apsr-to-replicate-everything` |
| Add a reanalysis repo | `vignette("reanalysis-studies")` |
| Use Stata (or bilingual R/Stata) | `vignette("stata-replications")` |
| Browse and run in the browser | `vignette("shiny-app")` |
| Audit the whole registry | `vignette("audit")` |
| Install dependencies on a server | `vignette("maintainer-setup")` |
| Package-backed studies | `vignette("package-replication-checklist")` |

Live app: [shiny2.wzb.eu/ipi/replicate/](https://shiny2.wzb.eu/ipi/replicate/). Registry: [github.com/replicate-anything/registry](https://github.com/replicate-anything/registry).
